---
title: 'The pbc trial'
output:
  pdf_document: default
  html_notebook: default
---

# Kaplan Meier survival time
```{r}
fit.KM=survfit(Surv(timeYears, event) ~ 1, data = dat.full)
##fit
plot(fit.KM)
title(main="Overall survival time (using Kaplan-Meier estimator)")
abline(v=9.3, col=2, lty=2)
abline(h=0.5, col=2, lty=2)
summary(fit.KM, time=9.3)
```

# trt
```{r}
survdiff(Surv(timeYears, event) ~ trt, data = dat.full)
survdiff(Surv(timeYears, event) ~ trt, data = subset(dat.full, dat.full$trt!="control"))
survdiff(Surv(timeYears, event) ~ trt, data = subset(dat.full, dat.full$trt!="placebo"))
survdiff(Surv(timeYears, event) ~ trt, data = subset(dat.full, dat.full$trt!="D-penicillamine"))
fit=survfit(Surv(timeYears, event) ~ trt, data = dat.full)
plot(fit, col=1:3)
legend("bottomleft", lty = 1, col = 1:3, legend = names(fit$strata))
title(main="survival time per treatment (Kaplan-Meier estimator)")

plot(fit, fun="cloglog", col = 1:8)

remove("fit")
```

# sex
```{r}
table.sex=table(dat.full$sex, exclude=NULL)
table.sex
barplot(table.sex)

survdiff(Surv(timeYears, event) ~ sex, data = dat.full)
fit.sex=survfit(Surv(timeYears, event) ~ sex, data = dat.full)
plot(fit.sex, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.sex$strata))

plot(fit.sex, fun="cloglog", col = 1:8)

remove(list=c("table.sex", "fit.sex"))
```

# ascites
```{r}
table.ascites=table(dat.full$ascites, exclude=NULL)

table.ascites
barplot(table.ascites)

table(dat.full[dat.full$trial,]$ascites, exclude=NULL)

survdiff(Surv(timeYears, event) ~ ascites, data = dat.full)

fit.ascites=survfit(Surv(timeYears, event) ~ ascites, data = dat.full)
plot(fit.ascites, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.ascites$strata))

plot(fit.ascites, fun="cloglog", col = 1:8)

remove(list=c("table.ascites", "fit.ascites"))
```

#hepato
```{r}
table.hepato=table(dat.full[dat.full$trial,]$hepato, exclude=NULL)
table.hepato
barplot(table.hepato)

survdiff(Surv(timeYears, event) ~ hepato, data = dat.full[dat.full$trial,])

fit.hepato=survfit(Surv(timeYears, event) ~ hepato, data = dat.full[dat.full$trial,])
plot(fit.hepato, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.hepato$strata))
title(main="survival time per hepatomegaly (Kaplan-Meier estimator)")

plot(fit.hepato, fun="cloglog", col = 1:8)

remove(list=c("table.hepato", "fit.hepato"))
```

#spiders
```{r}
table.spiders=table(dat.full[dat.full$trial,]$spiders, exclude=NULL)
table.spiders
barplot(table.spiders)

survdiff(Surv(timeYears, event) ~ spiders, data = dat.full[dat.full$trial,])
survdiff(Surv(timeYears, event) ~ spiders+strata(sex), data = dat.full[dat.full$trial,])

fit.spiders=survfit(Surv(timeYears, event) ~ spiders, data = dat.full[dat.full$trial,])
plot(fit.spiders, col = 1:2)
legend("bottomleft", lty = 1, col = 1:2, legend = names(fit.spiders$strata))
title(main='survival time per "spiders" (Kaplan-Meier estimator)')

plot(fit.spiders, fun="cloglog", col = 1:8)

fit.spiders=survfit(Surv(timeYears, event) ~ sex+spiders, data = dat.full[dat.full$trial,])
plot(fit.spiders, col = 1:4)
legend("bottomleft", lty = 1, col = 1:4, legend = names(fit.spiders$strata))
title(main='survival time per sex and "spiders"(Kaplan-Meier estimator)')

remove(list=c("table.spiders", "fit.spiders"))
```

#edema
```{r}
table.edema=table(dat.full$edema, exclude=NULL)
table.edema
barplot(table.edema)

survdiff(Surv(timeYears, event) ~ edema, data = dat.full)

fit.edema=survfit(Surv(timeYears, event) ~ edema, data = dat.full)
plot(fit.edema, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.edema$strata))

plot(fit.edema, fun="cloglog", col = 1:8)

remove(list=c("table.edema", "fit.edema"))
```

#stage
```{r}
table.stage=table(dat.full$stage, exclude=NULL)
table.stage
barplot(table.stage)

table(dat.full[dat.full$trial,]$stage, exclude=NULL)

survdiff(Surv(timeYears, event) ~ stage, data = dat.full)

# subset(dat.full, dat.full$trt!="control")
fit.stage=survfit(Surv(timeYears, event) ~ stage, data = dat.full)
plot(fit.stage, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.stage$strata))
title(main="Survival Time Per Stage (Kaplan-Meier Estimator)")

plot(fit.stage, fun="cloglog", col = 1:8)

remove(list=c("fit.stage", "table.stage"))
```

# age
```{r}
hist(dat.full$age)

dat <- within(dat.full, {
ageCat <- cut(age, breaks = c(0, 40, 50, 60, Inf))
})
table(dat$ageCat)

sum(is.na(dat.full[dat.full$trial,]$age))

survdiff(Surv(timeYears, event) ~ ageCat, data = dat)

fit.ageCat=survfit(Surv(timeYears, event) ~ ageCat, data = dat)
plot(fit.ageCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.ageCat$strata))

plot(fit.ageCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.ageCat"))
```

# bili

# Elevated bilirubin levels (>2.5-3 mg/dL) cause jaundice and can be classified into different anatomical sites of pathology: prehepatic (increased bilirubin production), hepatic (liver dysfunction), or posthepatic (duct obstruction).
```{r}
hist(dat.full$bili)

# c(0, 1, 2, 4, 8, 16, Inf))
dat <- within(dat.full, {
biliCat <- cut(bili, breaks = c(0, 1, 2, 3, 4, Inf))
})
table(dat$biliCat)

survdiff(Surv(timeYears, event) ~ biliCat, data = dat)

fit.biliCat=survfit(Surv(timeYears, event) ~ biliCat, data = dat)
plot(fit.biliCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.biliCat$strata))

plot(fit.biliCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.biliCat"))
```

# chol (healthy : less than 200 mg/dL)
```{r}
hist(dat.full$chol)

dat <- within(dat.full, {
cholCat <- cut(chol, breaks = c(0, 200, 250, 300, 400, Inf))
})
table(dat$cholCat)

survdiff(Surv(timeYears, event) ~ cholCat, data = dat)

fit.cholCat=survfit(Surv(timeYears, event) ~ cholCat, data = dat)
plot(fit.cholCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.cholCat$strata))

plot(fit.cholCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.cholCat"))
```

# albumin (normal range is 3.5 to 5.5 g/dL)

# Conditions associated with "high" levels of albumin: dehydration.
# Conditions associated with "low" levels of albumin are:
# - Ascites,
# - Liver disease ( hepatitis or cirrhosis)
# - Malnutrition
```{r}
hist(dat.full$albumin)

dat <- within(dat.full, {
albuminCat <- cut(albumin, breaks = c(0, 3, 3.5, 4, Inf))
})
table(dat$albuminCat)

survdiff(Surv(timeYears, event) ~ albuminCat, data = dat)

fit.albuminCat=survfit(Surv(timeYears, event) ~ albuminCat, data = dat)
plot(fit.albuminCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.albuminCat$strata))

plot(fit.albuminCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.albuminCat"))
```

# copper
# high value of total copper may indicate poor excretion secondary to underdeveloped biliary systems
```{r}
hist(dat.full$copper)

dat <- within(dat.full, {
copperCat <- cut(copper, breaks = c(0, 50, 100, 200, Inf))
})
table(dat$copperCat)

survdiff(Surv(timeYears, event) ~ copperCat, data = dat)

fit.copperCat=survfit(Surv(timeYears, event) ~ copperCat, data = dat)
plot(fit.copperCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.copperCat$strata))

plot(fit.copperCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.copperCat"))
```

# alk.phos
# man            : 53 to 128 U/L
# woman 20 to 50 : 42 to 98 U/L
# woman 60+      : 53 to 141 U/L

```{r}
hist(dat.full$alk.phos)

#
# c(0, 1000, 1500, 2000, 2500, Inf)
dat <- within(dat.full, {
alk.phosCat <- cut(alk.phos, breaks = c(0, 1000, 2000, Inf))
})
table(dat$alk.phosCat)
# dat <- subset(dat, dat$alk.phosCat!= "(2e+03,Inf]")
nrow(dat)
survdiff(Surv(timeYears, event) ~ alk.phosCat, data = dat)

fit.alk.phosCat=survfit(Surv(timeYears, event) ~ alk.phosCat, data = dat)
plot(fit.alk.phosCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.alk.phosCat$strata))

plot(fit.alk.phosCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.alk.phosCat"))
```

# ast
# Males: 6-34 IU/L
# Females: 8-40 IU/L

```{r}
hist(dat.full$ast)

# c(0, 80, 160, 240, Inf)
dat <- within(dat.full, {
astCat <- cut(ast, breaks = c(0, 80, 120, 160, Inf))
})
table(dat$astCat)
survdiff(Surv(timeYears, event) ~ astCat, data = dat)

fit.astCat=survfit(Surv(timeYears, event) ~ astCat, data = dat)
plot(fit.astCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.astCat$strata))

plot(fit.astCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.astCat"))
```

# trig
# normal :less than 150 mg per deciliter (mg/dL)

```{r}
hist(dat.full$trig)

# 
dat <- within(dat.full, {
trigCat <- cut(trig, breaks = c(0, 100, 200, Inf))
})
table(dat$trigCat)
survdiff(Surv(timeYears, event) ~ trigCat, data = dat)

fit.trigCat=survfit(Surv(timeYears, event) ~ trigCat, data = dat)
plot(fit.trigCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.trigCat$strata))

plot(fit.trigCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.trigCat"))
```

# platelet
# normal range : 150-450 billion/L
```{r}
hist(dat.full$platelet)

# c()
dat <- within(dat.full, {
plateletCat <- cut(platelet, breaks = c(0, 200, 300, Inf))
})
table(dat$plateletCat)
survdiff(Surv(timeYears, event) ~ plateletCat, data = dat)

fit.plateletCat=survfit(Surv(timeYears, event) ~ plateletCat, data = dat)
plot(fit.plateletCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.plateletCat$strata))

plot(fit.plateletCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.plateletCat"))
```

# protime
# Prothrombin time (PT) ?
#   man normal range is 9.6-11.8 seconds
# woman normal range is 9.5-11.3 seconds
```{r}
hist(dat.full$protime)

# c(0, 10, 10.5, 11, 12, Inf)
dat <- within(dat.full, {
protimeCat <- cut(protime, breaks = c(0, 10, 11, 12, Inf))
})
table(dat$protimeCat)
survdiff(Surv(timeYears, event) ~ protimeCat, data = dat)

fit.protimeCat=survfit(Surv(timeYears, event) ~ protimeCat, data = dat)
plot(fit.protimeCat, col = 1:8)
legend("bottomleft", lty = 1, col = 1:8, legend = names(fit.protimeCat$strata))

plot(fit.protimeCat, fun="cloglog", col = 1:8)

remove(list=c("dat", "fit.protimeCat"))
```