---
title: 'The pbc trial'
output:
  pdf_document: default
  html_notebook: default
---


```{r}
paste(names(dat.trial), collapse="+")
```

# full model (build from output above):
# trt+age+sex+ascites+hepato+spiders+edema+bili+albumin+alk.phos+ast+protime+stage+chol_fxd+trig_fxd+copper_fxd+platelet_fxd

```{r}
M.full=coxph(Surv(timeYears, event) ~ trt+age+sex+ascites+hepato+spiders+edema+
               bili+albumin+alk.phos+ast+protime+stage+chol_fxd+trig_fxd+
               copper_fxd+platelet_fxd, data = dat.trial)
summary(M.full)
```

# use a step approach to build a reduced model
```{r}
M.AIC <- step (M.full, trace=0)
summary(M.AIC)

fits <- list(M.full = M.full, M.AIC = M.AIC)
sapply(fits, AIC)
```

# check the proportionality of hazard for each variable
# H0: proportionality of hazard

```{r}
cox.zph(M.AIC)
```


```{r}
M.AIC2=coxph(Surv(timeYears, event) ~ age + edema + bili + albumin + ast + protime + copper_fxd + stage, data = dat.trial)
summary(M.AIC2)

fits <- list(M.full = M.full, M.AIC = M.AIC, M.AIC2 = M.AIC2)
sapply(fits, AIC)
```


```{r}
table(dat.trial$stage)
plot(
  survfit(Surv(timeYears, event) ~ stage, data = dat.trial),
  fun="cloglog", col = 1:4)
title(main='survival time per stage (log-log scale)')
```

# age + edema + bili + albumin + ast + protime + copper_fxd + strata(stage)

```{r}
M.AIC_StrStage=coxph(Surv(timeYears, event) ~ age + edema + bili + albumin + ast + protime + copper_fxd + strata(stage), data = dat.trial)
summary(M.AIC_StrStage)

fits <- list(M.full = M.full, M.AIC = M.AIC, M.AIC_StrStage=M.AIC_StrStage)
sapply(fits, AIC)
```

# age + bili + albumin + ast + protime + copper_fxd + strata(stage, edema)

```{r}
table(dat.trial$stage, dat.trial$edema)

M.AIC_StrStageEdema=coxph(Surv(timeYears, event) ~ age + bili + albumin + ast + protime + copper_fxd + strata(stage, edema), data = dat.trial)
summary(M.AIC_StrStageEdema)

fits <- list(M.AIC = M.AIC, M.AIC_StrStage=M.AIC_StrStage, M.AIC_StrStageEdema=M.AIC_StrStageEdema)
sapply(fits, AIC)
```


# age + edema + bili + albumin + ast + protime + copper_fxd + strata(stage, sex)

```{r}
table(dat.trial$stage, dat.trial$sex)

M.AIC_StrStageSex=coxph(Surv(timeYears, event) ~ age + edema + bili + albumin + ast + protime + copper_fxd + strata(stage, sex), data = dat.trial)
summary(M.AIC_StrStageSex)

fits <- list(M.AIC = M.AIC, M.AIC_StrStage=M.AIC_StrStage, M.AIC_StrStageSex=M.AIC_StrStageSex)
sapply(fits, AIC)
```


```{r}
hist(dat.trial$bili)

dat.trial$biliCat <- cut(dat.trial$bili, breaks=quantile(dat.trial$bili), include.lowest = TRUE)

table(dat.trial$biliCat)

fit.biliCat <- survfit(Surv(timeYears, event) ~ biliCat, data = dat.trial)
plot(fit.biliCat, fun="cloglog", col = 1:4)
legend("topleft", lty = 1, col = 1:4, legend = names(fit.biliCat$strata))
title(main='survival time per biliCat (log-log scale)')
```


```{r}
M.AIC_StrStage_BiliCat=coxph(Surv(timeYears, event) ~ age + edema + biliCat + albumin + ast + protime + copper_fxd + strata(stage), data = dat.trial)
summary(M.AIC_StrStage_BiliCat)

fits <- list(M.AIC = M.AIC, M.AIC_StrStage=M.AIC_StrStage, M.AIC_StrStage_BiliCat=M.AIC_StrStage_BiliCat)

sapply(fits, AIC)
```

```{r}
biliQuant=quantile(dat.trial$bili)
dat.trial$biliCat2 <- cut(dat.trial$bili, breaks=biliQuant[-2], include.lowest = TRUE)

table(dat.trial$biliCat2)
fit.biliCat2 <- survfit(Surv(timeYears, event) ~ biliCat2, data = dat.trial)
plot(fit.biliCat2, fun="cloglog", col = 1:4)
legend("topleft", lty = 1, col = 1:3, legend = names(fit.biliCat2$strata))
title(main='survival time per biliCat2 (log-log scale)')
```


```{r}
M.AIC_StrStage_BiliCat2=coxph(Surv(timeYears, event) ~ age + edema + biliCat2 + albumin + ast + protime + copper_fxd + strata(stage), data = dat.trial)
summary(M.AIC_StrStage_BiliCat2)

cox.zph(M.AIC_StrStage_BiliCat2)

fits <- list(M.AIC = M.AIC, M.AIC_StrStage=M.AIC_StrStage, M.AIC_StrStage_BiliCat2=M.AIC_StrStage_BiliCat2)

sapply(fits, AIC)
```


```{r}
M.AIC_StrStage_BiliLog=coxph(Surv(timeYears, event) ~ age + edema + log(bili) + albumin + ast + protime + copper_fxd + strata(stage), data = dat.trial)
summary(M.AIC_StrStage_BiliLog)

cox.zph(M.AIC_StrStage_BiliLog)

fits <- list(M.AIC = M.AIC, M.AIC_StrStage=M.AIC_StrStage, M.AIC_StrStage_BiliLog=M.AIC_StrStage_BiliLog)

sapply(fits, AIC)
```
