```{r}
library(survival)
library(survivalROC)
data(mayo)

dat.full <- pbc
dat.trial <- dat.full[!is.na(dat.full$trt),]
dat.trial$event <- 0 + (dat.trial$status == 2)

dat.sort <- dat.trial[order(dat.trial$event, dat.trial$time),]
mayo.sort <- mayo[order(mayo$censor, mayo$time),]
```

```{r}
dat.trial$mayoscore4 = NA
dat.trial$mayoscore5 = NA

for(ind in seq_along(dat.trial[,1])) {
  time <- (mayo$time == dat.trial[ind,]$time)
  event <- (mayo$censor == dat.trial[ind,]$event)
  if (sum(time & event) != 1) {
    indm <- which(time & event)[1]
  }
  else {
    indm <- which(time & event)
  }
  dat.trial[ind,]$mayoscore4 = mayo[indm,]$mayoscore4
  dat.trial[ind,]$mayoscore5 = mayo[indm,]$mayoscore5
}

sum(is.na(dat.trial$mayoscore4))
sum(is.na(dat.trial$mayoscore5))

rm(list=c("mayo", "mayo.sort", "dat.full", "dat.sort"))

dat.trial_1 = subset(dat.trial,dat.trial$stage==1)
dat.trial_2 = subset(dat.trial,dat.trial$stage==2)
dat.trial_3 = subset(dat.trial,dat.trial$stage==3)
dat.trial_4 = subset(dat.trial,dat.trial$stage==4)

```



```{r}
hist(dat.trial$mayoscore4)
```

```{r}
hist(dat.trial$mayoscore5)
```

The survival rate is rather high, about 40% at 12 year

```{r}
fit=survfit(Surv(time / 365.25, event) ~ 1, data = dat.trial)
plot(fit)
```

survival at 5 years
```{r}
summary(fit, time=5)
```

```{r}
summary(coxph(Surv(time,event)~mayoscore5, data=dat.trial))
```
```{r}
summary(coxph(Surv(time,event)~mayoscore4, data=dat.trial))
```
as both p values are very small, we may use the conformance to pick the best model

the ROC curves can be used to compared models if based on a single continuous covariate
predict.time is the time point at which the prediction is made (see function help)

```{r}
dat.trial_p = dat.trial_3

ROC.4 <- survivalROC(Stime = dat.trial_p$time,
                     status = dat.trial_p$event,
                     marker = dat.trial_p$mayoscore4,
                     predict.time = 365.25 * 5,
                     method="KM")
ROC.5 <- survivalROC(Stime = dat.trial_p$time,
                     status = dat.trial_p$event,
                     marker = dat.trial_p$mayoscore5,
                     predict.time = 365.25 * 5,
                     method = "KM")
```

```{r}
ROC <- list(mayo4 = ROC.4, mayo5 = ROC.5)
sapply(ROC, "[[", "AUC")
```
compute True Positive and False Positive
```{r}
dfl <- lapply(ROC, function(x) with(x, data.frame(FP, TP)))
for(nm in names(dfl)) {
  dfl[[ nm ]]$marker <- nm
}
dat <- do.call(rbind, dfl)
```

plot AUC curve False Positive / True Positive
```{r}
library(ggplot2)
ggplot(dat, aes(FP, TP, color = marker)) +
  geom_line() +
  theme_bw(base_size = 9)
```
use mayo5 because it has the lowest AUC
10% value is an arbitrary value
```{r}
cutoff <- with(ROC$mayo4, min(cut.values[FP <= 0.1]))
```
## 7.511961

```{r}
dat.trial_p$prediction <-
  ifelse(dat.trial_p$mayoscore4 <= cutoff,
         "low_risk", "high_risk")
```

the curves are displayed in alphabetical order
the split is optimal at 5 years because this is the point that was selected

```{r}
fit.KM <- survfit(Surv(time/365, event) ~ prediction, data = dat.trial_p)
plot(fit.KM, col = c("red", "blue"))
```
```{r}
fit.KM
```

